Constraining the High Density Nuclear Symmetry Energy with Pions

### 2020-05-15 chapters/ch_experiment.tex
First draft before Bill for Experimental section

\begin{figure}
\includegraphics[width=\textwidth]{tpcPrinciple.pdf}
\label{fig:tpcPrinciple}
\caption{Operation principle of the TPC}
\end{figure}
%explain in breif the basics of a time projection chamber 
Time projection chambers are a class of detectors which reconstruct charged particles in all 3-dimensions. Here we will outline the physical principles involved in the TPC measurement and discuss the particular details of the TPC used in this thesis. Inside of at TPC the detector gas is housed in a field cage, which also sets up a constant electric field. As charged particles pass through the gas, electron-ion pairs from neutral gas molecules by passing tracks; the electrons are accelerated opposite to the electric field and the ions move in the opposite direction. Since the mean free path of the electrons inside the gas is very small, they quickly collide with other gas molecules, slowing down the electron, which is then accelerated again repeating the cycle. This microscopic behavior manifests into a constant drift velocity when averaged over several gas collisions. These electrons drift up towards a set of wire planes eventually reaching a set of high voltage anode wires where they accelerate under the high electric field, liberating more electron-ion pairs from the gas creating an avalanche. The avalanche electrons finally terminate either on the anode wire or the grounded pad readout plane,  while the ions from the avalanche move slowly away from the anode wires; creating a large signal which is distributed over the pad-plane where the charge and time information of the signal is measured by the  electronics.

Two of the 3 coordinates are determined from the charge distribution on the pad plane. The third dimension comes from projecting the electrons back in time, utilizing the known constant drift velocity $v_d$; the distance the electron has traveled , $d$, -- along the electric field direction-- is calculated as $d = v_d \cdot t$, where $t$ is the time it took for the electrons to reach the electronics. The radius of curvature is related to the magnetic rigidity, and therefore the momentum of each track. The energy loss  deposited ($\langle dE/dx\rangle$) is measured by the segmented charge sensitive pads on the pad-plane, which are connected to the readout electronics. Particle identification is achieved through since unique particles exist on unique rigidity and $\langle dE/dx\rangle$  lines, which will be discussed in latter sections. In this chapter we will discuss in more detail the process described above in the context of the specific TPC used in this thesis.

@@

The Samurai Pion-Reconstruction and Ion Tracker Time Projection Chamber (\spirit TPC) is a multi-wire proportional counter developed to measure pions and other light charge particles resulting from radioactive heavy ion collisions in fixed target experiments.  The TPC is built on an aluminum angle iron skeleton with thin aluminum sheet walls all around, in order to minimize neutron scattering and to allow for light charged particles to reach auxilliary detectors on the sides and downstream of the TPC. The \spirit TPC was developed to fit inside the SAMURAI dipole magnet used at the Rare Isotope Beam Factory (RIBF) at RIKEN in Wako-shi, Japan \cite{riken}; the dipole gap limited the vertical space of the TPC to around \SI{75}{\centi\metre}. More detail and specifications of the SAMURAI dipole magnet are given in \cite{samurai}.

A target mechanism allowed for up to 5 fixed targets to be mounted at anytime, with the ability to change targets on the outside of the TPC. The field cage contained the detector gas and set up the constant electric field, which was mounted to a large aluminum top plate, though electrically isolated by a lexan top perimeter ring with o-rings to provide as gas seal. The pad plane and wire plane structures are also mounted to the inside face of the top plate with the electronics being mounted on the outside face of the top plate; several aluminum ribs were also mounted to provide extra rigidity to the top plate, keeping it flat to within \SI{150}{\micro\metre}, as measured by a laser \cite{jon}. Holes on the top plate allowed for the readout of the individual charge sensitive pads on the pad plane, through surface mount pads which were connected through short cables to the electronics. The exploded drawing shown in Fig.~\ref{fig:tpcExplode} pictures all of the major internal components of the  of the \spirit TPC

@@

\subsection{Enclosure}
The skeleton of the enclosure is composed of a rigid aluminum angle-iron frame. All walls are constructed of a aluminum frame with thin sheet metal. All materials were made to be as thin to allow charged particles and neutrons to exit the TPC without scattering too much. This allows for a trigger to be created by placing detectors on the sides and downstream of the TPC. The enclosure itself is made to be gas tight with respects to the outside and the field cage. This is to allow for the possibility to run a different gas inside the enclosure than the field cage. Although, in this set of experiments we ran the same gas in the field cage and enclosure volume.

@@

The field cage was constructed from several panels of printed circuit boards (PCBs). The front of the field cage was made of two PCBs and each side was constructed of three PCBs supported by Lexan pieces. The epoxy in the common PCB substrate FR4 contains bromine which is not suitable for the long term operation of a TPC, as the bromine will eventually cause gain reduction of the wires \cite{tpcAging}. The halogen free material used was Cryogenic-G10.  The field cage is an isolated volume from the enclosure of the TPC for the option to run explosive gases such hydrogen in the field cage, where the enclosure can be filled with an insulating gas. While the risk of a high voltage spark was minimized using the voltage step down, the risk of sparking when using an explosive gas could be further minimized by isolating the detector volume from the enclosure volume thereby allowing you to run an insulating gas between the field cage while running the explosive gas inside the detector volume only. Instead of a downstream wall, a large thin exit window was constructed which consisted of a lexan frame bonded to a \SI{10}{\micro\metre} Kapton window with evaporated aluminum strips. The PCB boards were epoxied into the cathode which was constructed of an aluminum honeycomb laminate where two sheets of aluminum were bonded to a core of aluminum honeycomb structure providing a lightweight yet rigid structure. On the other end the boards were epoxied into an aluminum top perimeter which also served as the last ring in the TPC. Together with the cathode bottom, the field cage proved to be a rigid lightweight structure. A lexan ring containing o-rings was placed in-between the top perimeter piece and the top plate of the TPC. Screws with nylon washers, and collars, were used to mount the top perimeter, and the field cage, to the top plate. The top plate could be removed and rotated with the field cage on without damaging any internal components.

@@

The cathode is connected to the HV supply through a \SI{10}{\mega\ohm} resistor and has an effective capacitance to ground of \SI{4}{\nano\farad}, $C_{VSD}$. The cathode voltage $V_{cath}$ can be calculated from the schematic of the TPC system in Fig.~\ref{fig:TPC_schematic}, as
\begin{equation}
V_{cath} = \frac{V_{HV}}{ 1 + \frac{10}{ \left( (245 + R_p)^{-1} + 700^{-1} \right)^{-1} } },
\end{equation}
where 
\begin{equation}
R_p = \left( R_{TP1}^{-1} + R_{TP2}^{-1} \right)^{-1},
 \label{eq:Reff}
\end{equation}

@@

Figure~\ref{fig:FC_schematic} shows a cartoon schematic of the field cage walls. The field cage contains 50 inside copper strips and 49 outside copper strips. The strips on each board was connected to the adjacent strip on the next board; front, back, and side boards were connected by G10 corner pieces with conducting paint strips, creating rings around the whole field cage. The first inside strip is connected to the cathode, which is itself connected by an effective \SI{5}{\mega\ohm} resistor ($R$) to the first outside strip. The first outside strip is connected trough a via to the second inside strip, which continues on in this repeating fashion. The resistor chain creates a voltage divider in which each strip is separated by a constant difference voltage at a fixed distance, setting up a constant electric field. The last strip of the field cage is composed of a small inner strip  (\SI{1.5}{\milli\metre}) on the PCB board and the aluminum top perimeter piece (\SI{4.5}{\milli\metre}) giving an  effective thickness of \SI{6}{\milli\metre}, the same as the other strip widths. The top perimeter is connected to electrical ground through a \SI{20}{\mega\ohm} resistor ($R_{TP1}$) with the option to place an additional resistor ($R_{TP2}$) in parallel to tune the voltage of the top perimeter, as seen in Fig.~\ref{fig:FC_schematic}.

@@

\subsection{Voltage Step Down}
The gap between the field cage's cathode and the ground of the enclosure is quite small. To prevent electric breakdown in the gas between this gap a series of concentric copper rings safely stepped down the voltage to ground; where each ring was separated by a resistor. There were 8 concentric rings with a \SI{10}{\mega\ohm} resistors in between, creating a resistor chain which steps down the voltage each ring by approximately \SI{1000}{\volt} each time. The first ring is the same voltage as the cathode and the last ring is connected to ground. All together the total resistance of the resistor chain is \SI{700}{\mega\ohm}.

@@

There are three wire planes that are mounted underneath the pad-plane. The wire plane closest to the pad-plane (\SI{4}{\milli\metre}) are the anode wires. The next plane (\SI{12}{\milli\metre}) is the ground plane or frisch grid, and the last plane (\SI{14}{\milli\metre}) is the gating grid. The gating grid is the first plane that electrons meet as they drift upward from the field cage volume towards the anode plane. The gating grid is operated as a gate either allowing electrons and ions through, or blocking them entirely. The ground plane functions to shield the inside volume of the TPC from the high electric field surrounding the anode wires where the avalanche process of the electron takes place. The ground plane is the least interesting plane and is shorted to the enclosure ground by a BNC terminator on the outside of the TPC. If we would like to calibrate the electronics of the TPC we typically inject a pulser into one end of the ground plane and terminate the other end with a \SI{50}{\ohm} termination.

@@

In the open configuration the gating grid is transparent to electrons coming from the field cage volume and also allows for ions to move from the avalanche region into the TPC volume. Typically the gating grid is held in the closed configuration only opening when the data acquisition trigger criteria is met. This is to block the electrons which come from the un-reacted beam, which happens frequently, which if allowed to go to the anode wires, would quickly build up enormous amounts of positive ions and flood the volume of the field cage with space charge. We close the gating grid after reading out approximately one TPC volume -- about \SI{10}{\micro\second}  -- to also prevent the back-flow of ions from the avalanche region from that event. Since ions move with a velocity much slower than that of elections, the ions only move several \si{\micro\metre} in the time the gate is open; this allows for electrons to pass through while preventing the back-flow of ions into the FC volume.

Figure~\ref{fig:gg_onoff} shows a Garfield simulation of the gating grid in both the on and off configurations. In the on configuration, all the wires share the same average voltage, $V_{g.g.}$, which is optimized for the case of 100\% electron transparency. In the off configuration, the reference voltage $V_{g.g.}$ remains the same, but alternating wires get an offset voltage of $\pm \Delta V$, so that the electric field produced by the voltage difference $2\Delta V$ between wires is great enough to block incoming electrons. Opening the grid from this closed bi-polar mode is simply done by removing the offset voltage and allowing the two wires to short which equilibrates their charges, which is the state of the reference voltage.

@@

Both configurations of the gating grid were measured and simulated. To measure the electron transparency were all wires share the voltage $V_{avg}$, the anode wire was lowered to \SI{500}{\volt} and the beam was allowed to enter the field cage without any target put in. By lowering the voltage of the anode we could measure the large charge of the beam without saturating the electronics. The average charge deposited in the chamber could be measured as a function of $V_{avg}$; changing the top plane resistor appropriately according to Eq.~\ref{eq:TP_resistor}. Several runs were taken ranging from \SIrange{-198}{-40}{\volt}, with and without the magnetic field. Theoretically the most negative value represents 100\% electron transparency and was used as the reference run. The electron transparency, $T$, was defined as $T = \langle dE/dx\rangle/{\langle dE/dx\rangle}_{ref}$, where $\langle dE/dx\rangle_{ref}$ represents the average energy loss of the reference run. Figure~\ref{fig:ggAvgTrans} shows the measured transparency as a function of $V_{avg}$, as compared with the corresponding Garfield simulation. The average gating grid voltage used in the experiment was \SI{-171}{\volt} to ensure we were well within the 100\% transparency region.

@@

The anode wires are made of very thin Gold plated Tungsten wires, about \SI{20}{\micro\metre} in diameter. They are biased to high voltages around \SI{e3}{\volt} which creates a very high electric field very close to the anode wire. As the eletron drifts towards the anode wire, it gains kinetic energy knocking out more electron-ion pairs in the gas until it terminates on the anode wire or the pad plane. The amount of electrons produced depends on the anode wire voltage and the gas properties. The absolute gas gain was not experimentally measured but was calculated in a Garfield simulation. During the experiment the anode wires were biased to two different voltages. We will refer to the voltage \SI{1460}{\volt} as the ``high voltage" and \SI{1214}{\volt} as the ``low voltage". Only two sections were biased with the lower voltage setting due to concerns of a high current issue. Figure~\ref{fig:anodegain} shows the expected number of electrons distributions for electrons produced in a avalanche process of an electron creating a single avalanche. The distribution follows the expected Polya distribution and the MC data in the simulation was fitted with a Polya function \cite{blumrol}, described as,

@@

The wire plane and the cathode define the different electric field regions. The main drift region is defined by the space between the gating grid and cathode voltages (Region 1), with a small drift region being defined by the space between the ground grid and the gating grid (Region 2); the avalanche region is defined as between the ground grid and the anode grid. The voltage of the gating grid is set to the point that allows for 100\% transparency for electrons. It is possible that the electric field in Region 1 and Region 2 can be matched but not required. The voltage of the top perimeter is set by adjusting the value of the last resistor ($R_{TP2}$) to ensure the electric field is constant through Region 1. We can imagine Region 1 is split into two virtual volumes, one defined between the cathode and the top perimeter, and one defined between the top perimeter and the gating grid. The magnitude of the electric field in the region between the top perimeter and the cathode, $E_1$, is defined as,

\begin{equation}
E_1 = \frac{V_{g.g.} - V_{tp}}{ y_{g.g.} - y_{tp} },
\end{equation}
where  $V_{g.g.}$, $V_{tp}$ , $y_{g.g.}$, and $y_{tp}$ are the voltage and vertical y-position of the gating grid and top perimeter respectively. The magnitude of the electric field in the region between the top perimeter and the cathode, $E_2$, is defined as,

\begin{equation}
E_2 = \frac{V_{tp} - V_{cath}}{ y_{tp} - y_{cath} },
\end{equation}
where  $V_{tp}$, $V_{cath}$ , $y_{tp}$, and $y_{cath}$ are the voltage and vertical y-position of the top perimeter and cathode respectively. The condition for a smooth electric field across the volume is given by $E_1 = E_2$. Substituting Eq.~\ref{eq:FCstrip} for $V_{tp}$ ($n=50$), we can solve for the effective resistance of the top perimeter $R_p$ as,

@@

where the relevant vertical dimensions are $y_{g.g.} - y_{cath} = \SI{497.3}{\milli\metre}$ and $y_{tp} - y_{cath} = \SI{490}{\milli\metre}$. The value of $R_{TP2}$ can be calculated from Eq.~\ref{eq:Reff} and Eq.~\ref{eq:TP_resistor}.

\subsection{Pad Plane}
The pad-plane is a multi-layer circuit board which is segmented into \SI{11.5}{\milli\metre} x \SI{7.5}{\milli\metre} charge sensitive pads; arranged in an array of 108 x 112 pads in the x and z-directions respectively making 12096 pads in total. There is an insulating gap of \SI{0.5}{\milli\metre} on each side separating the pads. The effective area covered by the pads is \SI{1344}{\milli\metre} x \SI{864}{\milli\metre}. There is a via and trace coming from each pad to the opposite side of the pad plane and is arranged in a surface pads which may be readout by a surface mount SAMTEC connector. Figure~\ref{fig:padplane} shows the pad plane boards being glued to the top plate and the holes which allow for the readout of pads. The pads were gold plated for excellent electric conduction properties.

@@

Signals in the S$\pi$RIT TPC are amplified and digitized by the recently developed Generic Electronics for TPCs (GET) \cite{get}.  Short cables transmit the signals from the pads to through a circuit protection board called ZAP, to the inputs of the AGET chips mounted to the AsAd board as seen in Fig.~\ref{fig:getzap}. Each AGET chip services 64 pads (63 pads are connected in our case), contains a pre-amplifier, and a Switched Capacitor Array (SCA), with a maximum of 512 time buckets with an adjustable sampling frequency of 1 to 100 MHz. Four AGET chips are mounted on one AsAd (ASIC and ADC) motherboard. The gain of each AGET can be configured as 0.12, 0.24, 1.0, or 10 pC over the whole dynamic range, and the ADCs on each AsAd board provides 12 bit resolution. The peaking times of the shaping amplifiers can be set to 69, 117, 232, 501, 720, or 1014 ns. In this experiment, the gain was set to the highest setting, 0.12 pC, the peaking time 117 ns, and the sampling frequency 25 MHz (resulting in 40 ns time buckets).

@@

After each AsAd board has digitized the data it is sent to the Concentration Boards (CoBo). Each CoBo board can concentrate the data from 4 AsAd boards. The Multiplicity, Trigger, and Time module (MuTanT) \cite{get} provides the common trigger signal for all CoBo boards.  Each board sends the data to the DAQ server which writes to disk the data from each board, which was handled by two separete DAQ servers; saving to one common analysis server. The data could then be analyzed using the RIKEN High Performace Computing cluster or moved to the NSCL or MSU cluster for analysis. The Aget 2.0, asad 2.1, and cobo 1.0 firmware versions were used.

\section{Energy loss in material}
The average energy loss in a material can be described by the Bethe-Bloch equation,
\begin{equation}\label{eq:bb}
\frac{dE}{dx} = \frac{4\pi NZ^2e^4}{mc^2\beta^2} (ln \frac{2mc^2\beta^2\gamma^2}{I} - \beta^2).
\end{equation}
Where $N$ is the number density of electrons in the medium, $e$ the elementary charge, $mc^2$ is the rest mass of the electron, $Z$ is the charge of the traversing particle, $I$ is the mean excitation energy of the medium, and $\beta$ is the velocity of the particle. There is a large variation in energy loss around this mean value, with a long high energy loss tail.

>>>>

\begin{figure}
\includegraphics[width=\textwidth]{tpcPrinciple.pdf}
\label{fig:tpcPrinciple}
\caption{Operation principle of the TPC}
\end{figure}
%explain in breif the basics of a time projection chamber 
Time projection chambers are a class of detectors which reconstruct charged particles in all 3-dimensions. Here we will outline the physical principles involved in the TPC measurement and discuss the particular details of the TPC used in this thesis. Figure~\ref{fig:tpcPrinciple} depicts the inside of a TPC, specifically the field cage, which holds the detector gas and also sets up a constant electric field. As charged particles resulting from a heavy ion collision pass through the gas, electron-ion pairs are created from neutral gas molecules. The electrons are accelerated opposite to the electric field and the ions along with it. Since the mean free path of the electrons inside the gas is very small, they quickly collide with other gas molecules; slowing down, or even stopping the electron, which is then accelerated again, starting a repeating cycle of stop and go motion. This microscopic behavior manifests as a constant drift velocity when averaged over several gas collisions. The electrons drift up towards a set of wire planes eventually reaching a set of high voltage anode wires where they quickly accelerate in the presence of the high electric field, liberating more electron-ion pairs from the gas, creating an avalanche process to occur. The avalanche electrons eventually terminate either on the anode wire or the grounded pad readout plane,  while the ions from the avalanche move slowly away from the anode wires, creating a large signal which is distributed over the pad-plane. The charge and timing  information of these electrons are calculated from the induced signal measured by the  electronics.

Two of the 3 coordinates are determined from the 2-dimensional charge distribution on the pad-plane. The third dimension comes from projecting the electrons back in time, utilizing the known constant drift velocity $v_d$. The distance the electron has traveled , $d$, -- along the electric field direction-- is calculated as $d = v_d \cdot t$, where $t$ is the timing information of the signal in the electronics. The radius of curvature of the track is related to the magnetic rigidity, and therefore the momentum of each track. The energy loss deposited $\langle dE/dx\rangle$ is measured by the segmented charge sensitive pads on the pad-plane, in which each pad is connected to a channel in the electronics. Particle can be uniquely identified since  each particles exists on unique rigidity and $\langle dE/dx\rangle$  lines, which will be discussed in latter sections. In this chapter we will discuss in more detail the process described above in the context of the specific TPC used in this thesis.

@@

The Samurai Pion-Reconstruction and Ion Tracker Time Projection Chamber (\spirit TPC) is a multi-wire proportional counter developed to measure pions and other light charge particles resulting from radioactive heavy ion collisions in fixed target experiments.  The TPC is built on an aluminum angle iron skeleton with thin aluminum sheet walls in order to minimize neutron scattering and to allow for light charged particles to reach auxilliary detectors on the sides and downstream of the TPC. The \spirit TPC was developed to fit inside the SAMURAI dipole magnet used at the Rare Isotope Beam Factory (RIBF) at RIKEN in Wako-shi, Japan \cite{riken}. The dipole gap limited the vertical space of the TPC to around \SI{75}{\centi\metre}. More detail and specifications of the SAMURAI dipole magnet are given in \cite{samurai}.

A target mechanism allowed for up to 5 fixed targets to be mounted at anytime, with the ability to change targets on the outside of the TPC. The field cage contained the detector gas and set up the constant electric field, which was mounted to a large aluminum top plate, though electrically isolated by a lexan top perimeter ring with o-rings to provide as gas seal. The pad plane and wire plane structures are also mounted to the inside face of the top plate with the electronics being mounted on the outside face of the top plate. Several aluminum ribs were also mounted to provide extra rigidity to the top plate, keeping it flat to within \SI{150}{\micro\metre}, as measured by a precise laser measurment \cite{jon}. Holes on the top plate allowed for the readout of the individual charge sensitive pads on the pad plane, through surface mount pads which were connected through short cables to the electronics. The exploded drawing shown in Fig.~\ref{fig:tpcExplode} pictures all of the major internal components of the  of the \spirit TPC.

@@

\begin{comment}
\subsection{Enclosure}
The skeleton of the enclosure is composed of a rigid aluminum angle-iron frame. All walls are constructed of a aluminum frame with thin sheet metal. All materials were made to be as thin to allow charged particles and neutrons to exit the TPC without scattering too much. This allows for a trigger to be created by placing detectors on the sides and downstream of the TPC. The enclosure itself is made to be gas tight with respects to the outside and the field cage. This is to allow for the possibility to run a different gas inside the enclosure than the field cage. Although, in this set of experiments we ran the same gas in the field cage and enclosure volume. 
\end{comment}

@@

The field cage was constructed from several panels of printed circuit boards (PCBs). The front of the field cage was made of two PCBs and each side was constructed from three PCBs, supported by Lexan pieces. The common PCB substrate material FR4, contains a bromine epoxy which can outgas and absorb electrons degrating the signal \cite{tpcAging}. Therefore a halogen free material, Cryogenic-G10, was used for the board material. The field cage volume is isolated from the enclosure volume of the TPC for the option to run two separate  gasses, which may be useful when running explosive gasses such hydrogen. Instead of a downstream wall a, large, thin exit window was constructed, which consisted of a  \SI{10}{\micro\metre} Kapton window with evaporated aluminum strips, which was bonded to a removable Lexan frame. The PCB boards were epoxied, and screwed, into the cathode which was constructed from an aluminum honeycomb laminate, composed of two aluminum sheets bonded to an aluminum honeycomb core, providing a lightweight yet rigid structure. On the other end the boards were epoxied into an aluminum top perimeter which also served as the last ring in the TPC. Together with the cathode bottom, the field cage proved to be a rigid lightweight structure. A Lexan ring containing o-rings was placed in-between the top perimeter piece and the top plate of the TPC. Screws with nylon washers, and collars, were used to mount the top perimeter --and therefore the field cage-- to the top plate; it also provided electrical isolation from the top plate. In this way the top plate could be removed and rotated with the field cage attached on without damaging any internal components.

@@

Figure~\ref{TPC_schematic} shows the schematic of the effective resistances and capacitance of the TPC subsystem. The cathode is connected to the HV supply through a \SI{10}{\mega\ohm} resistor and has an effective capacitance to ground of \SI{4}{\nano\farad}, $C_{VSD}$. The cathode voltage $V_{cath}$ can be calculated as,
\begin{equation}
V_{cath} = \frac{V_{HV}}{ 1 + \frac{10}{ \left( (245 + R_p)^{-1} + 700^{-1} \right)^{-1} } },
\end{equation}
where 
\begin{equation}
R_p = \left( R_{TP1}^{-1} + R_{TP2}^{-1} \right)^{-1},
 \label{eq:Reff}
\end{equation}

@@

Figure~\ref{fig:FC_schematic} shows a schematic detailing the connections involved in the field cage walls. The field cage contains 50 inside copper strips and 49 outside copper strips. The strips are \SI{6}{\milli\metre} in width and spaced \SI{10}{\milli\metre} apart. The strips on each board was connected to the adjacent strips on the adjacent boards. The interfaces were the side pieces met the front boards or back window are connected by G-10 corner pieces with conducting paint strips, which makes the strips into continuous rings around the whole field cage. The first inside strip is connected to the cathode, which is itself connected by an effective \SI{5}{\mega\ohm} resistor ($R$) to the first outside strip. The first outside strip is connected through a via to the second inside strip. This pattern repeats until the last strip. The resistor chain creates a voltage divider in which each strip is separated by a constant difference voltage and a fixed distance, setting up a constant electric field. The last strip of the field cage is composed of a small inner strip  (\SI{1.5}{\milli\metre}) on the PCB board and the top perimeter piece (\SI{4.5}{\milli\metre}) giving an  effective thickness of \SI{6}{\milli\metre}, the same as the other strip widths. The top perimeter is connected to electrical ground through a \SI{20}{\mega\ohm} resistor ($R_{TP1}$) with the option to place an additional resistor ($R_{TP2}$) in parallel to tune the voltage of the top perimeter, as seen in Fig.~\ref{fig:FC_schematic}.

@@

\subsection{Voltage Step Down}
The gap between the cathode and the ground of the enclosure is quite small. To prevent electric breakdown in the gas between this gap a series of concentric copper rings safely stepped down the voltage to ground in a controlled manner, reducing the chance of electric breakdown and sparking. There were 8 concentric rings with a \SI{10}{\mega\ohm} resistors in between, creating a resistor chain which steps down the voltage each ring by approximately \SI{1000}{\volt} each time. The first ring is the same voltage as the cathode and the last ring is connected to ground. All together the total resistance of the resistor chain is \SI{700}{\mega\ohm}.

@@

There are three wire planes that are mounted underneath the pad-plane. The wire plane closest to the pad-plane (\SI{4}{\milli\metre}) are the anode wires. The next plane (\SI{8}{\milli\metre}) is the ground plane or frisch grid, and the last plane (\SI{14}{\milli\metre}) is the gating grid. The gating grid is the first plane that electrons meet as they drift upward from the field cage volume towards the anode plane. The gating grid is operated as a gate, either allowing electrons and ions through, or blocking them entirely. The ground plane functions to shield the inside volume of the TPC from the high electric field surrounding the anode wires. The ground plane is the least interesting plane and is held to ground by shorting the plane to the enclosure through  shorted BNC terminator on the outside of the TPC. We also use the ground plane to input a pulser which is used to spread the pulsed signal to all the pads on the TPC in order to calibrate the electronics of the TPC. This is done by replacing the shorted BNC with a \SI{50}{\ohm} termination and injecting the pulser on the other end.

@@

In the open configuration, the gating grid is transparent to electrons coming from the field cage volume and also allows for ions to move from the avalanche region into the TPC volume. Typically the gating grid is held in the closed configuration, only opening it when the data acquisition trigger criteria is met. By keeping it always closed the electrons which come from the un-reacted beam are blocked,  which if allowed to go to the anode wires, would quickly build up enormous amounts of positive ions, and would flood the volume of the field cage with space charge. We open the gating grid for about \SI{11}{\micro\second} which is more than the time it takes for the electrons to drift one TPC volume. After this we close the gating grid to prevent the back-flow of ions from the avalanche region from that event. Since ions move with a velocity much slower than that of electrons \cite{blumrol}, the ions only move several \si{\micro\metre} in the time the gate is open; this allows for electrons to pass through while preventing the back-flow of ions into the FC volume.

Figure~\ref{fig:gg_onoff} shows a Garfield simulation of the drift lines of electrons in both the on and off configurations. In the on configuration, all the wires share the same average voltage, $V_{g.g.}$, in which the optimal voltage is selected for the case of 100\% electron transparency. Figure~\ref{fig:wires_open} shows the electrons are allowed to drift completely through the gating grid all the way to terminate on the anode wires. In the off configuration, the reference voltage $V_{g.g.}$ remains the same, but alternating wires get an offset voltage of $\pm \Delta V$, so that the electric field produced by the voltage difference $2\Delta V$ between wires is great enough to block incoming electrons. Figure~\ref{fig:wires_closed} shows this case were the electron drift lines are fully blocked terminating on the more positive wires. Opening the grid from this closed bi-polar mode is simply done by removing the offset voltage and allowing the two wires to short which equilibrates their charges, which is the steady state of the average reference voltage for 100\% transparency.

@@

Both configurations of the gating grid were measured and simulated. To measure the electron transparency were all wires share the voltage $V_{avg}$, the anode wire was lowered to \SI{500}{\volt} and the beam was allowed to enter the field cage without any target put in. By lowering the voltage of the anode wires we could measure the large charge of the beam without saturating the electronics. The average charge deposited in the chamber could be measured as a function of $V_{avg}$; changing the top plane resistor appropriately according to Eq.~\ref{eq:TP_resistor}. Several runs were taken ranging from \SIrange{-198}{-40}{\volt}, with and without the magnetic field. Theoretically the most negative value represents 100\% electron transparency and was used as the reference run. The electron transparency, $T$, was defined as $T = \langle dE/dx\rangle/{\langle dE/dx\rangle}_{ref}$, where $\langle dE/dx\rangle_{ref}$ represents the average energy loss of the reference run. Figure~\ref{fig:ggAvgTrans} shows the measured transparency as a function of $V_{avg}$, as compared with the corresponding Garfield simulation. The average gating grid voltage used in the experiment was \SI{-171}{\volt} to ensure we were well within the 100\% transparency region.

@@

The anode wires are made of very thin Gold plated Tungsten wires, about \SI{20}{\micro\metre} in diameter. They are biased to high voltages around \SI{e3}{\volt} which creates a very high electric field very close to the anode wire. As the eletron drifts towards the anode wire, it gains enough kinetic energy to knock out more electron-ion pairs in the gas which also go on to do the same; until they terminate on the anode wires or the pad plane. The amount of electrons produced depends on the anode wire voltage and the gas properties. The absolute gas gain was not experimentally measured but was calculated in a Garfield simulation. For the experimental data pertaining to this thesis, the anode wires were biased to two different voltages. We will refer to the voltage \SI{1460}{\volt} as the ``high voltage" and \SI{1214}{\volt} as the ``low voltage". Only two sections were biased with the lower voltage setting due to a leakage around the end of the gating grid \cite{jon}. Figure~\ref{fig:anodegain} shows the electron distribution for the total number of electrons produced in a avalanche process created by a single electron. The distribution follows the expected Polya distribution, and the MC data in the simulation was fitted with a Polya function \cite{blumrol}, which can be expressed as,

@@

The main drift region is defined by the space between the gating grid and cathode voltages (Region 1), with a small drift region being defined by the space between the gating grid and the ground grid (Region 2), and the avalanche region is defined as between the ground grid and the anode grid. It is possible that the electric field in Region 1 and Region 2 can be matched but not required.  To ensure the electric field is continuous across Region 1, the voltage of the top perimeter  must be set properly by adjusting the value of the last resistor ($R_{TP2}$). We can imagine Region 1 is split into two virtual volumes, one defined as the volume between the cathode and the top-perimeter, and one defined between the top-perimeter and the gating-grid. The magnitude of the electric field in the region between the top-perimeter and the cathode, $E_1$, is defined as,

\begin{equation}
E_1 = \frac{V_{g.g.} - V_{tp}}{ y_{g.g.} - y_{tp} },
\end{equation}
where  $V_{g.g.}$, $V_{tp}$ , $y_{g.g.}$, and $y_{tp}$ are the voltages and vertical y-positions of the gating-grid and top-perimeter respectively. The y-position here refers to the center of the electrodes. The magnitude of the electric field in the region between the top-perimeter and the cathode, $E_2$, is defined as,

\begin{equation}
E_2 = \frac{V_{tp} - V_{cath}}{ y_{tp} - y_{cath} },
\end{equation}
where  $V_{tp}$, $V_{cath}$ , $y_{tp}$, and $y_{cath}$ are the voltage and vertical y-position of the top-perimeter and cathode respectively. The y-position of the cathode is defined as the face of the cathode. The condition for a smooth electric field across these two virtual volumes is defined as the solution to the equation $E_1 = E_2$. Substituting Eq.~\ref{eq:FCstrip} for $V_{tp}$ -- $n=50$ -- we can solve for the effective resistance of the top perimeter $R_p$ as,

@@

where the relevant vertical dimensions are $y_{g.g.} - y_{cath} = \SI{497.3}{\milli\metre}$ and $y_{tp} - y_{cath} = \SI{490}{\milli\metre}$. The value of $R_{TP2}$ can then be calculated from Eq.~\ref{eq:Reff}.

\subsection{Pad Plane}
The pad-plane is a multi-layer circuit board which is segmented into \SI{11.5}{\milli\metre} x \SI{7.5}{\milli\metre} charge sensitive pads; arranged in an array of 108 x 112 pads in the x and z-directions respectively, making 12096 pads in total. There is an insulating gap of \SI{0.5}{\milli\metre} on each side separating the pads so that the effective area covered by the pads is \SI{1344}{\milli\metre} x \SI{864}{\milli\metre}. There is a via and trace coming from each pad, through the board, to the opposite side of the pad plane and is arranged in a surface pads which is readout by a surface mount SAMTEC connector. Figure~\ref{fig:padplane} shows the pad plane boards being glued to the top plate and the holes which allow for the readout of pads. The pads were gold plated for excellent electric conduction properties.

@@

Signals in the S$\pi$RIT TPC are amplified and digitized by the recently developed Generic Electronics for TPCs (GET) \cite{get}.  Short cables transmit the signals from the surface mount connectors, through a circuit protection board called ZAP, to the inputs of the AGET chips which are mounted to the AsAd board as seen in Fig.~\ref{fig:getzap}. Each AGET chip services 64 pads (63 pads are connected in our case). Four AGET chips are mounted on one AsAd  motherboard. Figure~\ref{fig:aget} is the schematic of each  AGET chip which contains a charge sensitive pre-amplifier, several other stages of amplifiers, and a Switched Capacitor Array (SCA) with a maximum of 512 time buckets which operates in a circular readout buffer. The sampling frequency can be adjusted from 1 to \SI{100}{\mega\hertz}. The gain of each AGET can be configured as 0.12, 0.24, 1.0, or \SI{10}{\pico\coulomb} over the whole dynamic range, and the analog-to-digital converters (ADCs) on each AsAd board provides 12 bit resolution. The peaking times of the shaping amplifiers can be set to 69, 117, 232, 501, 720, or \SI{1014}{\nano\second}. In this experiment, the gain was set to the highest setting, 0.12 \si{\pico\coulomb}, the peaking time \SI{117}{\nano\second}, and the sampling frequency \SI{25}{\mega\hertz} (resulting in \SI{40}{\nano\second} time buckets).

@@

After each AsAd board has digitized the data it is sent to the Concentration Boards (CoBo). Each CoBo board can concentrate the data from 4 AsAd boards. The Multiplicity, Trigger, and Time module (MuTanT) \cite{get} provides the common trigger signal for all CoBo boards.  Each board sends the data to the DAQ server which writes to disk the data from each board, which was handled by two separete DAQ servers; saving to one common analysis server. The data could then be analyzed using the RIKEN High Performance Computing (HPC) cluster or moved to the NSCL or MSU cluster for analysis. The Aget 2.0, asad 2.1, and cobo 1.0 firmware versions were used in this analysis.
